---
title: "Project 2: STAT302Package Tutorial"
author: "Juan Hillon"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{STAT302Package Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE}
library(STAT302Package)
data("my_gapminder")
data("my_penguins")
```

## Introduction
This package demonstrates package building, using functions from previous labs. 
These functions are `my_t.test`, a t test function, `my_lm`, a linear model function, `my_knn_cv`, a k-nearest neighbors cross-validation function, and `my_rf_cv`, a rondom forest cross-validation function. I will be using the datasets `my_penguins` and `my_gapminder` to demonstrate these functions.

To install and load STAT302Package, use the following code:
```{r, eval=FALSE}
devtools::install_github("juanhillon/STAT302Package")
library(STAT302Package)
```

## Tutorial for `my_t.test`
To explain how to use `my_t.test`, I will be demonstrating three t-tests with a p-value cut-off of 0.05 using the `lifeExp` data from `my_gapminder` and explaining the results. This function returns a list with four elements. The first item represents the test statistic, which is equal to `(mean(x) - mu) / std_error`. The second item represent the degrees of freedom used in calculating the p-value, which is equal to `length(x) -` `. The third item returns the type of alternative test the user specified. The fourth item is the p-value of null hypothesis test.

The first t-test I will demonstrate is:
  \begin{align}
  H_0: \mu &= 60,\\
  H_a: \mu &\neq 60.
  \end{align}
This is an example of a two-sided t-test, with `mu` being equal to 60. The null hypothesis is that `mu` equals 60, and the alternative hypothesis is that mu is not equal to 60. Therefore we set up a `my_t.test` function that looks like this:
```{r}
my_t.test(x = my_gapminder$lifeExp, alternative = "two.sided", mu = 60)
```
This gives us a p-value of `r my_t.test(x = my_gapminder$lifeExp, alternative = "two.sided", mu = 60)[4]`. This is greater than the cut-off of 0.05, so we do not reject the null hypothesis.


The second t-test I will demonstrate is:
  \begin{align}
  H_0: \mu &= 60,\\
  H_a: \mu &< 60.
  \end{align}
This is an example of a t-test where the alternative is less than `mu`, with `mu` being equal to 60. The null hypothesis is that `mu` equals 60, and the alternative hypothesis is that mu is less than 60. Therefore we set up a `my_t.test` function that looks like this:
```{r}
my_t.test(x = my_gapminder$lifeExp, alternative = "less", mu = 60)
```
This gives us a p-value of `r my_t.test(x = my_gapminder$lifeExp, alternative = "less", mu = 60)[4]`. This is less than the cut-off of 0.05, so we reject the null hypothesis.

The third t-test I will demonstrate is:
  \begin{align}
  H_0: \mu &= 60,\\
  H_a: \mu &> 60.
  \end{align}
This is an example of a t-test where the alternative is greater than `mu`, with `mu` being equal to 60. The null hypothesis is that `mu` equals 60, and the alternative hypothesis is that mu is greater than 60. Therefore we set up a `my_t.test` function that looks like this:
```{r}
my_t.test(x = my_gapminder$lifeExp, alternative = "greater", mu = 60)
```
This gives us a p-value of `r my_t.test(x = my_gapminder$lifeExp, alternative = "greater", mu = 60)[4]`. This is greater than the cut-off of 0.05, so we do not reject the null hypothesis.
